Towards a standardized evaluation of imputation methodology

### 2022-07-26 Manuscript/Towards_a_standardized_evaluation_of_imputation_methodology.tex
Revisions version shared with Gerko

[TODO; describe problems with data generation and interpretation.]

- If the source of the data is model-based, the correlations in the data may be so high that there is no effective missingness due to the predictive power in the other variables, and the data source may be unfair if the data generating model is equal to one of the imputation models under evaluation.

- With a design-based simulation, the data source is unlikely to have no missingness and the choice of initial missing data method may propagate into the evaluations.

@@

[TODO: paraphrase ``be clear which mechanisms are of interest, how they are generated, if they are suitably generated".]

- Missingness mechanisms can be spurious (spurious MAR when correlations are too low, or when missingness proportion is too low). In reporting, the functional form or type of missingness is often omitted. As advice, MCAR can be a very informative mechnism for reference and MNAR may be useful for practice.

- Missingness proportions can be misinterpreted.

- Missingness pattern may have unexpected consequences (e.g., univariately generated multivariate missingness). In interpretation, watch out for generalizations from unintended special cases.

@@

Even the terminology on missingness generation can be confusing. Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry? In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation. The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion of 25\% in both variables, or one completely observed variable and one with 50\% missingness).

@@

Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables. It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways. Extrapolating simulation results from a specific missing data pattern to more intricate empirical missingness patterns could lead to sub-optimal advice in practice. For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may conclude that list-wise deletion outperforms their imputation method(s). Such a conclusion would, unfortunately, only translate to data that adhere to the same special case while biasing inferences in other cases \citep[][\S 2.7]{buur18}. Another example of a missingness pattern that may inadvertently impose assumptions on generalizations is a monotone pattern \citep[i.e. a pattern with uniformly increasing missingness proportions along variables;][]{litt20}. If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration to produce valid results in all other situations. A final example is multivariate missingness which is generated using step-wise univariate missingness induction. The resulting incomplete data may then not have the desired statistical properties \citep[e.g. skewness;][]{ampute}. This complicates any definitive conclusions about imputation method performance in relation to the data-generating mechanism.

Missingness mechanisms are usually assumed to be random (MAR; see Table \ref{table:mech} for definitions). However, the missingness mechanisms that are induced in simulation studies do not necessarily match any true missingness generating mechanisms in incomplete empirical data, nor the mechanisms that are typically assumed when imputing such data. On the one hand, there is MCAR missingness, which is unlikely to be the true missingness generating mechanism, but would yield insightful evaluations in a simulation setting. On the other hand, there is MNAR, which is arguably a quite reasonable source of many real-world missingness scenarios but is generally ignored except for those evaluations that are specifically targeted at non-ignorable applications. A disconnect between induced missingness, real missingness, and assumed missingness may result in simulation studies that are not as informative as they could be.

MCAR is a necessary simulation condition for evaluating the performance of imputation procedures, yet often omitted. Under MCAR, the statistical properties of the observed data given the missing data are known. Any imputation routine that cannot at least mimic the performance of the observed data inference should be deemed inefficient in the scope of the simulation. If an imputation method is not able to solve the problem (i.e. yield valid inference) under MCAR, the statistical properties of the procedure are not sound. Sadly, the straightforward case of MCAR is often neglected from simulation studies and focus is drawn to the evaluation of MAR mechanisms only. Alternatively, some studies limit their evaluations to MCAR mechanisms only, which in our view may be far too simplistic.

@@

Inducing a MAR mechanism presents the simulator with another choice, namely the type or functional form of the missingness model. Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model \citep{scho18}. This may influence the performance of some (but not necessarily all) imputation routines. For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single MAR mechanism.

Finally, although one cannot definitively verify if the missingness is random--after all, for every MNAR mechanism there is a MAR mechanism with equal fit \citep{molenberghs2008every}--it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios. As stated, this mechanism is not usually included in evaluations.

@@

Missingness should be induced according to several sets of missing data conditions. We encourage simulators to consider different missingness mechanisms (including missingness types), and different missingness patterns (including missingness proportions). Although not every missingness mechanism is realistically assumed in practice, they can all offer valuable insights as simulation condition.

First, one should always consider MCAR missingness (see Table \ref{table:mech} for definitions). This mechanism should be used as a reference condition. Every imputation method must be able to yield valid imputations under MCAR, both in terms of distributional characteristics as well as statistical inference.

Next, missing data should be induced conform a model that is dependent on the observed data (i.e. a MAR mechanism). Arguably, MAR is the most often-assumed mechanism in practice, and should thus be evaluated carefully. A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}. If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly}. It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness. The effects of different types of MAR mechanisms are described in \citet{scho18}.

Third, a non-ignorable or MNAR mechanism might fall outside many studies' scope, yet could yield informative insights for daily practice. Even if an imputation method is specifically developed with ignorable missingness in mind, chances are that the method will be applied to non-ignorable empirical missingness at some point in time. It would, therefore, be wise to include MNAR missingness in the simulation study just in case. Alternatively, a sensitivity analysis may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.

In addition to varying the missingness mechanisms, each simulated mechanism should be combined with different missingness patterns. Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process. Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data. The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness. If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption. Alternatively, the validity of assumptions becomes increasingly important when the missingness increases. Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions. We, therefore, propose to evaluate imputation methodology under several missingness proportions to amulate a realistic range in severity of the missingness problem. Depending on how the missingness mechanism interacts with the simulated data, higher missingness proportions may yield biased or invalid completed data inferences. The missingness proportions should be considered carefully.

% \begin{table}[htb]
% \begin{center}
% \caption{Suggested univariate missingness proportions. [TODO: change this to 'more than one missingness proportion, depending on the context' and/or add references for the given proportions (e.g. about  that 30\% is a good range for MI).]}
% \label{table:prop}
% \begin{tabular}{lll}
% \hline
%       & Reasoning for suggested missingness proportion \\
% \hline  
% 10\%  & Depending on the size of the data, this percentage can be considered as a lower \\
%       & bound of realistic evaluation. Anything less than 10\% may be of little influence on \\
%       & the true data inference. Performance of a missing data method should at least be \\
%       & acceptable for most missing data problems. \\ \noalign{\smallskip} % & \\
% 25\%  & This is a fair amount of missingness and will, depending on the observed data \\
%       & information, have a noticeable influence on the completed data inference. When \\
%       & compared to the condition with 10\% missingness, the inference obtained under \\
%       & 25\% missingness should be less certain (i.e. confidence/credibility interval width \\
%       & should increase), but estimates should still be properly covered and the statistical \\
%       & properties of the missing data method should be sound. In practice, at least to our \\
%       & at least to our experience in social sciences and official statistics, 25\% \\
%       & univariate missingness can easily be considered as a realistic missingness percentage. \\ \noalign{\smallskip} % & \\
% 50\%  & Performance under 50\% percent simulated missingness will most likely be impacted \\ 
%       & severely. Depending on how the missingness mechanism interacts with the simulated \\
%       & data, some imputation techniques may yield estimates that are under-covered such that \\
%       & the completed data inference should not be deemed valid anymore. If a method yields \\
%       & acceptable inference under 50\% MAR missingness, we can determine that the statistical \\
%       & properties of the imputation methodology are sound. \\
% \hline
% \end{tabular}
% \end{center}
% \end{table}

@@

\subsection{Evaluation}

The imputation methods under evaluation should be applied to the generated incomplete data, after which an analysis model may be fitted to estimate the estimand(s). In addition, the simulator should always perform complete case analysis. We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating imputation performance. Complete case analysis may, therefore, serve as a benchmark method against which imputation performance should be evaluated. Moreover, pairing complete case analysis with MCAR and MAR mechanisms allow the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.

The evaluation criteria used to assess imputation performance vary from one simulator to another. This is not surprising as people from different fields could have a different focus on the problem at hand. There are, however, some overarching issues with assessing imputation method performance. In the first place, diagnostic evaluation of the generated imputations is often left out of simulation study results. Problems with the imputation-generating process (e.g., an iterative imputation algorithm) may then be overlooked. In the second place, there are pitfalls in the evaluation of the estimates that are obtained by fitting the analysis model (i.e. the complete data model used to estimate the comparative truth). The choice of performance measures might inadvertently influence simulation results.

Many contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, to generate imputations. As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated. Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{buur18} is neither efficient nor failproof. As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method. Problems with producing stable imputations is not an exclusive quality iterative imputation algorithms. Any method may run into failures of the imputation-generating process and subsequently lack results (e.g. due to over-parameterization errors).

Omitting further evaluation of the imputation-generating process may yield sub-optimal imputations. The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models. When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}. [TODO: change this to being able to embed the analysis procedure and the imputation model within a Bayesian model; the nonresponse mdel may be misspecified while the imputation procedure is congenial.]The congeniality of all imputation models should be assessed, but methods to assess the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}. This makes its assessment an infeasible endeavor in many simulation studies. Moreover, the performance of imputation procedures on distributional properties is often ignored too. Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers. For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large. While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny. Under many circumstances, imputation methods may be realistically expected to preserve both marginal and conditional distributions with respect to the comparative truth. Imputation methods that fail to do so, should not be considered general-purpose methods (e.g., mean imputation).

The second source of problems with evaluating simulation outcomes is the choice of performance measures. Performance measures should express how well the obtained estimates approximate the comparative truth, but not every metric is suited for the evaluation of imputation methodology. For example, using measures of predictive accuracy or the RMSE (root mean squared error) as the evaluation criterion may increase the rate of false positives \citep[][\S 2.6]{buur18}. And, under certain simulation conditions, an otherwise useful metric such as relative bias can distort the interpretation of results (i.e., a negligible absolute bias may yield an infinite relative bias if the true value of the estimand is zero). The usefulness of evaluation criteria may thus depend on the simulation goal and estimand.

@@

Every imputation workflow should contain an evaluation of the obtained imputations. Even though inspecting each imputation may be labor-intensive due to the number of imputations generated in a simulation study, we highly recommend to consider the following aspects:

@@

@@

Imputation method performance should be quantified using appropriate measures. It depends on the specifics of each simulation study which performance measures would be most suitable. In general, we recommend to evaluate at least the following points:

(i) The estimates should preferably be unbiased. Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results. Unbiased estimates under MCAR are a minimum requirement for any imputation method.

We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values \citep[][\S 2.5]{buur18}. However, if a study is aimed at obtaining predictions or classification and inferential validity is not of interest, the RMSE of the predictions may yield valuable information about the methods' efficiency in terms of both accuracy and precision. [TODO: make this an optional bullet point for prediction only: out of sample mean squared prediction error can be compared between methods.]

>>>>

[TODO; describe problems with data generation and interpretation: If the source of the data is model-based, the correlations in the data may be so high that there is no effective missingness due to the predictive power in the other variables, and the data source may be unfair if the data generating model is equal to one of the imputation models under evaluation. With a design-based simulation, the data source is unlikely to have no missingness and the choice of initial missing data method may propagate into the evaluations.]

@@

[TODO: Make clear in this section that missingness mechanisms can be spurious (spurious MAR when correlations are too low, or when missingness proportion is too low). In reporting, the functional form or type of missingness is often omitted. As advice, MCAR can be a very informative mechnism for reference and MNAR may be useful for practice. Missingness proportions can be misinterpreted. Missingness pattern may have unexpected consequences (e.g., univariately generated multivariate missingness). In interpretation, watch out for generalizations from unintended special cases.]

@@

Even the terminology on missingness generation can be confusing. Does a missingness proportion of 50\% mean that half of the entries in an incomplete dataset are missing, or that half of the rows have at least one missing entry? In this paper, we will henceforth refer to the latter as the proportion of incomplete cases, and keep the term `missingness proportion' restricted to the variable-by-variable interpretation. The distinction between the two concepts, however, is not always clear in the literature, and convoluting the terms may lead to incorrect generalizations because they rarely mean the same (e.g., a proportion of 50\% incomplete cases in bivariate data could translate to a missingness proportion of 25\% in both variables, or one completely observed variable and one with 50\% missingness). Simulators should therefore be explicit in their description of the missing data generating model fitting the aim of their simulations.

@@

% Another aspect of missingness patterns that may misguide evaluations is the complexity of the patterns across variables. It is hardly reasonable to imagine empirical data with only one incomplete variable, yet some simulation studies rely on univariate missingness patterns anyways.

Extrapolating simulation results from a specific missing data generating model to more intricate (empirical) missingness could lead to sub-optimal advice in practice. For example, if a simulator induces missingness in the outcome variable of their analysis model exclusively, they may inadvertedly induce missingness according to the special case descibed by \citet[][p. \S 2.7]{buur18}. Under this specific missingness model, list-wise deletion may outperform any imputation method. Such a conclusion would, unfortunately, only translate to data that adhere to the same special case, while biasing inferences in other cases. % Another example of a missingness pattern that may inadvertently impose assumptions on generalizations is a monotone pattern \citep[i.e. a pattern with uniformly increasinmissingness proportions along variables;][]{litt20}. If a simulator investigates iterative imputation methods under monotone missingness, they might believe that the evaluated imputation algorithms converge instantly, while these methods actually require iteration to produce valid results in all other situations. 
Another example of potential side-effects of missingness generation may occur when multivariate missingness is generated using step-wise univariate missingness induction. The resulting incomplete data may then not have the desired statistical properties \citep[e.g. skewness;][]{ampute}. This complicates any definitive conclusions about imputation method performance in relation to the data generation parameters.

% Missingness mechanisms are usually assumed to be random (MAR; see Table \ref{table:mech} for definitions). However, the missingness mechanisms that are induced in simulation studies do not necessarily match any true missingness generating mechanisms in incomplete empirical data, nor the mechanisms that are typically assumed when imputing such data. On the one hand, there is MCAR missingness, which is unlikely to be the true missingness generating mechanism, but would yield insightful evaluations in a simulation setting. On the other hand, there is MNAR, which is arguably a quite reasonable source of many real-world missingness scenarios but is generally ignored except for those evaluations that are specifically targeted at non-ignorable applications. A disconnect between induced missingness, real missingness, and assumed missingness may result in simulation studies that are not as informative as they could be.

Missingness should be induced according to several sets of missing data conditions. We encourage simulators to consider different missingness patterns and mechanisms. Missingness mechanisms were first defined in Rubin's seminal work \citep{rubi76}. For more recent considerations see e.g. \citet{seam13, meal15, dore18, more18, moha21, scho21}. Although not every missingness mechanism is realistically assumed in practice, they can all offer valuable insights as simulation condition. A disconnect between induced missingness, real missingness, and assumed missingness may result in simulation studies that are not as informative as they could be.

[TODO: add that focusing on a single mechanism (even if the mechanism suits the study aim) is wasteful?] Truly random missingness across all data entries (`missing completely at random' or MCAR under Rubin's definition) may be considered a necessary simulation condition for the evaluation of imputation procedures, since the statistical properties of the observed data given the missing data are known. Any imputation routine that cannot at least mimic the performance of the observed data inference should be deemed inefficient in the scope of the simulation. If an imputation method is not able to solve the problem (i.e. yield valid inference) under MCAR, the statistical properties of the procedure are not universally sound. Sadly, the straightforward case of MCAR is often neglected from simulation studies, while it offers an informative reference condition in many simulation setups.

[TODO: maybe add definitions of missingness mechanisms back in???] A straightforward technique for inducing univariate MAR missingness is described in \citet[][\S 3.2.4]{buur18}, generalizations to multivariate MAR missingness can be found in \citet{ampute}. If the missingness is to be induced in longitudinal data autoregressive MAR models can be useful \citep[see e.g.][model 2 and model 3]{shara2015randomly}. It is advisable to investigate varying shapes of MAR missingness to achieve a more realistic indication of the robustness of the imputation performance across the range of random missingness. The effects of different types of MAR mechanisms are described in \citet{scho18}. Inducing a MAR mechanism presents the simulator with a choice, namely the type or functional form of the missingness model. Given the simulated data distributions, one random missingness model may be far more disastrous to the observed information than another model \citep{scho18}. This may influence the performance of some (but not necessarily all) imputation routines. For example, inference from hot-deck techniques such as predictive mean matching \citep{little1988missing, rubin1986statistical} may be more severely impacted by large amounts of one-tailed missingness than inference from parametric techniques. It would be a shame to overlook such results due to the focus on a single functional form of the missingness-generating model.

@@

A non-ignorable or MNAR mechanism might fall outside many studies' scope, yet could yield informative insights for daily practice. Even if an imputation method is specifically developed with ignorable missingness in mind, chances are that the method will be applied to non-ignorable empirical missingness at some point in time. After all, for every MNAR mechanism there is a MAR mechanism with equal fit and one cannot definitively verify that empirical missingness is random \citep{molenberghs2008every}. Therefore, it can be argued that MNAR is the more likely mechanism for real-life missingness scenarios. It may be wise to include MNAR missingness in the simulation study just in case. Alternatively, a sensitivity analysis may provide an indication of the validity of the obtained inference, given that the assumed missingness mechanism is suspected to be invalid \citep[see e.g.][part 5]{molenberghs2014handbook}.

@@

% First, one should always consider MCAR missingness (see Table \ref{table:mech} for definitions). This mechanism should be used as a reference condition. Every imputation method must be able to yield valid imputations under MCAR, both in terms of distributional characteristics as well as statistical inference.
% 
% Next, missing data should be induced conform a model that is dependent on the observed data (i.e. a MAR mechanism). Arguably, MAR is the most often-assumed mechanism in practice, and should thus be evaluated carefully.

In addition to varying the missingness mechanisms, each simulated mechanism should be combined with different missingness patterns. Remember that missingness is only ignorable under MAR when the parameter of the data is distinct and a-priori independent from the parameter of the missing data process. Under MAR missingness we assume that we may use the observed data to make inferences about the joint (observed and unobserved) data. The dependency of the procedure on the assumption under which we obtain inference is only influenced by the amount of missingness. If there is no missingness--or if there is no data, for that matter--the inference does not depend on the assumption. Alternatively, the validity of assumptions becomes increasingly important when the missingness increases. Since we control the MAR mechanism, the assumption under which we may solve the missing data problem should hold and it is only fair to assess performance under stringent missingness conditions. We, therefore, propose to evaluate imputation methodology under several missingness proportions to amulate a realistic range in severity of the missingness problem. Depending on how the missingness mechanism interacts with the simulated data, higher missingness proportions may yield biased or invalid completed data inferences. The missingness proportions should therefore be considered carefully.

[TODO: paraphrase ``be clear which mechanisms are of interest, how they are generated, if they are suitably generated" as conclusion to this section.]
% \begin{table}[htb]
% \begin{center}
% \caption{Suggested univariate missingness proportions. [TODO: change this to 'more than one missingness proportion, depending on the context' and/or add references for the given proportions (e.g. about  that 30\% is a good range for MI).]}
% \label{table:prop}
% \begin{tabular}{lll}
% \hline
%       & Reasoning for suggested missingness proportion \\
% \hline  
% 10\%  & Depending on the size of the data, this percentage can be considered as a lower \\
%       & bound of realistic evaluation. Anything less than 10\% may be of little influence on \\
%       & the true data inference. Performance of a missing data method should at least be \\
%       & acceptable for most missing data problems. \\ \noalign{\smallskip} % & \\
% 25\%  & This is a fair amount of missingness and will, depending on the observed data \\
%       & information, have a noticeable influence on the completed data inference. When \\
%       & compared to the condition with 10\% missingness, the inference obtained under \\
%       & 25\% missingness should be less certain (i.e. confidence/credibility interval width \\
%       & should increase), but estimates should still be properly covered and the statistical \\
%       & properties of the missing data method should be sound. In practice, at least to our \\
%       & at least to our experience in social sciences and official statistics, 25\% \\
%       & univariate missingness can easily be considered as a realistic missingness percentage. \\ \noalign{\smallskip} % & \\
% 50\%  & Performance under 50\% percent simulated missingness will most likely be impacted \\ 
%       & severely. Depending on how the missingness mechanism interacts with the simulated \\
%       & data, some imputation techniques may yield estimates that are under-covered such that \\
%       & the completed data inference should not be deemed valid anymore. If a method yields \\
%       & acceptable inference under 50\% MAR missingness, we can determine that the statistical \\
%       & properties of the imputation methodology are sound. \\
% \hline
% \end{tabular}
% \end{center}
% \end{table}

@@

\subsection{Performance evaluation}

[TODO: make this section less stringent and focus more on the suitability of the performancxe evaluation in the context of the study aim.]

The evaluation criteria used to assess imputation performance vary from one simulator to another. This is not surprising as people from different fields could have a different focus on the problem at hand. There are, however, some overarching issues with assessing imputation method performance. In the first place, there are pitfalls in the evaluation of the estimates that are obtained by fitting the analysis model after imputation (e.g. focusing on one performance measure over another, may yield different conclusions about imputation efficacy). In the second place, diagnostic evaluation of the generated imputations is often left out of simulation study results. Identifying problems with the imputation-generating process (e.g., an iterative imputation algorithm) may offer explanations for under-performance of imputation methods. Such evaluations, however, are typically omitted and valuable insights into the imputation method(s) may then be overlooked. %(i.e. the complete-data model used to estimate the comparative truth). % The choice of performance measures might inadvertently influence simulation results.

At minimum, imputation method performance should be quantified using appropriate measures. It depends on the specifics of each simulation study which performance measures would be most suitable. We recommend to evaluate at least the following points:

(i) The estimates should preferably be unbiased. Note that the way bias is calculated should be carefully chosen and described, since this can greatly influence the interpretation of the results. In most cases, unbiased estimates may be expected under a MCAR missingness mechanism.

(iv) Predictive accuracy may be quantified using the root mean squared error (RMSE). We do not generally recommend the use of the RMSE as evaluation criterion, because this metric does not account for the inherent uncertainty of the missing values \citep[][\S 2.5]{buur18}. However, if a study is aimed at obtaining predictions or classification and inferential validity is not of interest, the RMSE of the predictions may yield valuable information about the methods' efficiency in terms of both accuracy and precision. [TODO: make this an optional bullet point for prediction only: out of sample mean squared prediction error can be compared between methods.]

[TODO: intrgrate this paragraph into the above!] Performance measures should express how well the obtained estimates approximate the comparative truth, but not every metric is suited for the evaluation of imputation methodology. For example, using measures of predictive accuracy or the RMSE (root mean squared error) as the evaluation criterion may increase the rate of false positives \citep[][\S 2.6]{buur18}. And, under certain simulation conditions, an otherwise useful metric such as relative bias can distort the interpretation of results (i.e., a negligible absolute bias may yield an infinite relative bias if the true value of the estimand is zero). The usefulness of evaluation criteria may thus depend on the simulation goal and estimand.

@@

Next to imputation method performance, the imputation-generating process may be evaluated. Omitting further evaluation of the imputation-generating process may yield sub-optimal imputations. The evaluation of simulated results relies on the relation between missing data models, imputation models, and analysis models. When an imputation model is able to capture the essence of the true non-response mechanism relative to the analysis model, the models are said to be congenial \citep{meng94}. [TODO: change this to being able to embed the analysis procedure and the imputation model within a Bayesian model; the nonresponse mdel may be misspecified while the imputation procedure is congenial.] The congeniality of all imputation models should be assessed, but methods to assess the suitability of imputation models and diagnose misfit often rely on visual inspection of the imputations \citep[see e.g.][]{abayomi2008diagnostics, bond16}. This makes its assessment an infeasible endeavor in many simulation studies. Moreover, the performance of imputation procedures on distributional properties is often ignored too. Even though the estimates on the analysis level may be justified, some methods can yield imputations that may seem completely invalid to applied researchers. For example, one could very accurately estimate average human height by filling in negative values and values that are unrealistically large. While the obtained inference could still be valid under such imputations, the plausibility of the imputed values given the observed data should be under scrutiny. Under many circumstances, imputation methods may be realistically expected to preserve both marginal and conditional distributions with respect to the comparative truth. Imputation methods that fail to do so, should not be considered general-purpose methods (e.g., mean imputation).

[TODO: add segway to this paragraph.] Many contemporary imputation techniques rely on iterative algorithms, such as the Gibbs sampler, to generate imputations. As with any iterative algorithm--but especially with imputation algorithms that are critically considered to be possibly incompatible Gibbs samplers \citep[PIGS,][]{li2012imputing}--algorithmic convergence should be carefully evaluated. Unfortunately, there is no universal quantitative method to diagnose non-convergence in iterative imputation algorithms \citep{zhu15, ober21} and the alternative \citep[visual inspection of the imputation algorithm;][]{buur18} is neither efficient nor failproof. As a result, imputation algorithms may be terminated before reaching a stable state, which could yield sub-optimal imputations and under-estimated performance of the method. Problems with producing stable imputations is not an exclusive quality iterative imputation algorithms. Any method may run into failures of the imputation-generating process and subsequently lack results (e.g. due to over-parameterization errors).

[TODO: add segway to this paragraph and/or intrgrate into text above!] Every imputation workflow should contain an evaluation of the obtained imputations. Even though inspecting each imputation may be labor-intensive due to the number of imputations generated in a simulation study, we highly recommend to consider the following aspects:

@@

[TODO: add segway to reference method paragraph and add that performance on the complete data ofefrs an upper limit for performance.] In addition, the simulator should always perform complete case analysis. We know the theoretical properties of complete case analysis, which makes the technique useful as a point of departure when evaluating imputation performance. Complete case analysis may, therefore, serve as a benchmark method against which imputation performance should be evaluated. Moreover, pairing complete case analysis with MCAR and MAR mechanisms allow the simulator to evaluate the validity of the missing data generation process under the chosen analysis model.

@@
